### 2025-04-18 base-rates.qmd
20250418 - rookie performance

```{r}
library("petersenlab")
```

@@

@@

@@

>>>>

```{r}
library("petersenlab")
library("psych")
library("tidyverse")
```

### Load Data {#sec-baseRatesLoadData}

```{r}
#| eval: false
#| include: false

load(file = "./data/nfl_depthCharts.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
```

```{r}
load(file = "./data/player_stats_weekly.RData")
load(file = "./data/player_stats_seasonal.RData")
```

We created the `player_stats_weekly.RData` and `player_stats_seasonal.RData` objects in @sec-calculatePlayerAge.

@@

We examine the base rates of rookie performance compared to performance of non-rookies.
Rookies who play tend to be better players who were drafted in early rounds of the National Football League (NFL) draft.
Thus, to more fairly compare rookies and non-rookies, we examine only players who were drafted in the first two rounds of the NFL Draft.

```{r}
rookiesDraftedEarly <- player_stats_seasonal %>% 
  filter(years_of_experience == 0) %>% 
  filter(draftround %in% c(1,2)) %>% 
  arrange(player_id, season) %>% 
  group_by(player_id) %>% 
  slice_head(n = 1) %>% 
  ungroup() %>% 
  arrange(player_display_name)

nonRookiesDraftedEarly <- player_stats_seasonal %>% 
  filter(years_of_experience > 0) %>% 
  filter(draftround %in% c(1,2)) %>% 
  arrange(player_display_name)
```

[EXAMINE FOR ROOKIES WHO WERE DRAFTED IN FIRST TWO ROUNDS OF NFL DRAFT]

[ALSO, FOR A SECONDARY ANALYSIS, SUBSET TO PLAYERS WHO PLAYED A FULL SEASON]

@@

```{r}
nonRookiesDraftedEarly %>% 
  filter(position == "QB") %>%
  select(fantasyPoints) %>% 
  psych::describe()
```

```{r}
rookiesDraftedEarly %>% 
  filter(position == "QB") %>%
  select(fantasyPoints) %>% 
  psych::describe()

rookiesDraftedEarly %>% 
  filter(position == "QB") %>%
  select(fantasyPoints) %>% 
  hist()
```

Among Quarterbacks who played a full season:

```{r}
nonRookiesDraftedEarly %>% 
  filter(position == "QB") %>%
  filter(games >= 16) %>% 
  select(fantasyPoints) %>% 
  psych::describe()
```

```{r}
rookiesDraftedEarly %>% 
  filter(position == "QB") %>%
  filter(games >= 16) %>% 
  select(fantasyPoints) %>% 
  psych::describe()
```

@@

```{r}
nonRookiesDraftedEarly %>% 
  filter(position == "RB") %>%
  select(fantasyPoints) %>% 
  psych::describe()

nonRookiesDraftedEarly %>% 
  filter(position == "RB") %>%
  select(rushing_tds) %>% 
  psych::describe()
```

```{r}
rookiesDraftedEarly %>% 
  filter(position == "RB") %>%
  select(fantasyPoints) %>% 
  psych::describe()

rookiesDraftedEarly %>% 
  filter(position == "RB") %>%
  select(fantasyPoints) %>% 
  hist()

rookiesDraftedEarly %>% 
  filter(position == "RB") %>%
  select(rushing_tds) %>% 
  psych::describe()
```

Among Running Backs who played a full season:

```{r}
nonRookiesDraftedEarly %>% 
  filter(position == "RB") %>%
  filter(games >= 16) %>% 
  select(fantasyPoints) %>% 
  psych::describe()

nonRookiesDraftedEarly %>% 
  filter(position == "RB") %>%
  filter(games >= 16) %>% 
  select(rushing_tds) %>% 
  psych::describe()
```

```{r}
rookiesDraftedEarly %>% 
  filter(position == "RB") %>%
  filter(games >= 16) %>% 
  select(fantasyPoints) %>% 
  psych::describe()

rookiesDraftedEarly %>% 
  filter(position == "RB") %>%
  filter(games >= 16) %>% 
  select(rushing_tds) %>% 
  psych::describe()
```

### Wide Receivers {#sec-baseRateRookiePerformanceWRs}

```{r}
nonRookiesDraftedEarly %>% 
  filter(position == "WR") %>%
  select(fantasyPoints) %>% 
  psych::describe()

nonRookiesDraftedEarly %>% 
  filter(position == "WR") %>%
  select(receiving_tds) %>% 
  psych::describe()
```

```{r}
rookiesDraftedEarly %>% 
  filter(position == "WR") %>%
  select(fantasyPoints) %>% 
  psych::describe()

rookiesDraftedEarly %>% 
  filter(position == "WR") %>%
  select(fantasyPoints) %>% 
  hist()

rookiesDraftedEarly %>% 
  filter(position == "WR") %>%
  select(receiving_tds) %>% 
  psych::describe()
```

Among Wide Receivers who played a full season:

```{r}
nonRookiesDraftedEarly %>% 
  filter(position == "WR") %>%
  filter(games >= 16) %>% 
  select(fantasyPoints) %>% 
  psych::describe()

nonRookiesDraftedEarly %>% 
  filter(position == "WR") %>%
  filter(games >= 16) %>% 
  select(receiving_tds) %>% 
  psych::describe()
```

```{r}
rookiesDraftedEarly %>% 
  filter(position == "WR") %>%
  filter(games >= 16) %>% 
  select(fantasyPoints) %>% 
  psych::describe()

rookiesDraftedEarly %>% 
  filter(position == "WR") %>%
  filter(games >= 16) %>% 
  select(receiving_tds) %>% 
  psych::describe()
```

### Tight Ends {#sec-baseRateRookiePerformanceTEs}

```{r}
nonRookiesDraftedEarly %>% 
  filter(position == "TE") %>%
  select(fantasyPoints) %>% 
  psych::describe()

nonRookiesDraftedEarly %>% 
  filter(position == "TE") %>%
  select(receiving_tds) %>% 
  psych::describe()
```

```{r}
rookiesDraftedEarly %>% 
  filter(position == "TE") %>%
  select(fantasyPoints) %>% 
  psych::describe()

rookiesDraftedEarly %>% 
  filter(position == "TE") %>%
  select(fantasyPoints) %>% 
  hist()

rookiesDraftedEarly %>% 
  filter(position == "TE") %>%
  select(receiving_tds) %>% 
  psych::describe()
```

Among Tight Ends who played a full season:

```{r}
nonRookiesDraftedEarly %>% 
  filter(position == "TE") %>%
  filter(games >= 16) %>% 
  select(fantasyPoints) %>% 
  psych::describe()

nonRookiesDraftedEarly %>% 
  filter(position == "TE") %>%
  filter(games >= 16) %>% 
  select(receiving_tds) %>% 
  psych::describe()
```

```{r}
rookiesDraftedEarly %>% 
  filter(position == "TE") %>%
  filter(games >= 16) %>% 
  select(fantasyPoints) %>% 
  psych::describe()

rookiesDraftedEarly %>% 
  filter(position == "TE") %>%
  filter(games >= 16) %>% 
  select(receiving_tds) %>% 
  psych::describe()
```
